\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-09-13 redaction/JEGeo/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement .tex

\textbf{JE REPORTE ICI LA VERSION DANS LE MAIN TEXT DE LA VERSION SOUMISE. IL FAUT RESUMER.}

@@

\paragraph{Obtaining the ``pure'' transport costs component series} - The empirical strategy to find the fitted transport cost measure (i.e. extracting from trade composition effect) can be described as a two-stage process.
First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects.
For the estimated ad-valorem component, we estimate the following equation:
\begin{eqnarray}
\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}

\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country and sector fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country, product and year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows for the relevant year.
As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g. the series of $\gamma_t$ differs from one estimation to the other).
Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}

%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
\begin{eqnarray}
\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
\end{eqnarray}
As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
Both equations are estimated by transport mode.\smallskip

In this exercise, we are interested in isolating the change in the time dimension of each transport cost component.
This constitutes the second stage of our procedure.
As for the ad-valorem component defined in $[1;+\infty]$, from the estimation of Equation (\ref{eq:compeffects_mult}), we build the variable $\Gamma^{adv}_t$, for each year $t\geq 1974$, according to:
\begin{equation}
\Gamma^{adv}_t = 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1} \label{eq:comp_effects_adv}
\end{equation}

\noindent with $\bar{\tau}_{1974} = \exp(\delta +\sum_i \alpha_i +\sum_s \beta_s)$ the mean ad-valorem transport cost in 1974 (see details in Appendix \ref{app:comp-effects}).
In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).

As for the additive cost defined in $[0;+\infty]$, we build the variable $\Gamma^{add}_t$, the reference year being 1974 (ie, with $\gamma_{1974}=0$) according to:
\begin{equation}
\Gamma^{add}_t = 100.\exp(\gamma_t) \label{eq:comp_effects_add}
\end{equation}

\noindent As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.

\paragraph{Obtaining the unfitted measures} - The objective is to find the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:

-------IL FAUT INTRODUIRE LES PRIX FOB. Sinon, les coûts additifs ne sont pas modifiés de l’inflation -----

@@

We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e. composition effects excluded).
As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974.
Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
\smallskip

\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}.
Recall that $\widehat{\tau}^{adv}_t-1$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the fas price.
We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$

@@

$$\widehat{\tau}^{pure, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$

@@

with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost.

We then deduce the fitted value of the overall cost according to:
$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \widehat{t}^{pure}_t$$

@@

\paragraph{For the ad-valorem component} - Consider first the multiplicative transport cost component.
Rewriting Equation (\ref{eq:compeffects_mult}) by taking the exponential, we get:

\begin{equation*}
\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbbm{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbbm{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
\end{equation*}

Based on this equation, we deduce after estimation that:

From this, we obtain the following recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$.
Given that $\tau >1$, we can rewrite to get the percentage change between year 1974 and any year $t>1974$:
\begin{equation*}
\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}-1}{\widehat{\tau}_{is74}-1} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)-1}{\widehat{\tau}_{is74}-1}
\end{equation*}

As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the cost observed in 1974 and the time trend.
At this stage though, it remains specific to a product-origin country pair.
Next step is to build the index $\Gamma^{adv}_t$ such that:
\begin{equation}
 \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}  \label{eq:tcadv_compoeffect}
\end{equation}
\noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.

\paragraph{For the additive component} - After estimating Equation (\ref{eq:compeffects_add}), we can rebuild the additive component according to:
\[\left\{
  \begin{array}{lcl}
\text{For the year 1974:}&&\widehat{t}_{is74}=  \delta + \alpha_i+ \beta_s, \\
\text{For any year}~t> 1974:&&\widehat{t}_{ist}=\left(\delta + \alpha_i+ \beta_s\right).\exp(\gamma_t)
\end{array}
\right.\]

From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.
Given the constraint $t>0$, we then obtain the percentage change from 1974 from:

\begin{equation}
\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t) \label{eq:indice_add}
\end{equation}

\noindent Note that it is independent of the product-origin country pair. We can thus rewrite the time-trend series for the additive transport cost component as:
\begin{equation}
\Gamma^{add}_t  = 100\exp(\gamma_t) \label{eq:tcadd_compoeffect}
\end{equation}

>>>>

@@

\paragraph{Obtaining the ``pure'' transport costs component series} - The empirical strategy to find the fitted transport cost measure (i.e. extracting from trade composition effect) can be described as a two-stage process.
First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects. In plain words, we measure how these costs have changed over time by blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample).\smallskip

For the estimated ad-valorem component, we estimate the following equation:
\begin{eqnarray}
\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}

\noindent where $\mathbbm{1}_i$ and $\mathbbm{1}_{s(k)}$ represent country and sector fixed effects.\footnote{Throughout the exercise, we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country, product and year dummies.}  Equation (\ref{eq:compeffects_mult}) is estimated using OLS, with a weighting scheme based on the value of each flow in the total value of flows for the relevant year.

In this exercise, we are interested in isolating the change in the time dimension of each transport cost component. This constitutes the second stage of our procedure. Based on Equation (\ref{eq:compeffects_mult}), we deduce after estimation that:

From this, we obtain the following recursive link: $\widehat{\tau}_{ist} = \widehat{\tau}_{is74}\exp(\gamma_t)$. Given the constraint $\tau>0$, we rewrite to get the percentage change between year 1974 and any year $t>1974$:
\begin{equation*}
\Gamma_{ist} = 100.\frac{\widehat{\tau}_{ist}}{\widehat{\tau}_{is74}} = 100.\frac{\widehat{\tau}_{is74}\exp(\gamma_t)}{\widehat{\tau}_{is74}} = 100\exp(\gamma_t)
\end{equation*}

As such, the index of transport costs in year $t$ (relative to the reference year 1974) $\Gamma_{ist} $  only depends on the cost observed in 1974 and the time trend. Note that it is independent of the product-origin country pair, such that we can rewrite:
\begin{equation}
 \Gamma^{adv}_t= 100\exp(\gamma^{adv}_t)  \label{eq:tcadv_compoeffect}
\end{equation}

\noindent where the subscript $^{adv}$ clarifies that this time fixed effect is specific to the ad-valorem component. \medskip

\textbf{REPRENDRE AVEC GUILLAUME. ON ESTIME TAU>1 OU TAU>0? CA CHANGE, CF ANCIENNE VERSION EN ITALIQUE}

%At this stage though, it remains specific to a product-origin country pair.
%Next step is to build the index $\Gamma^{adv}_t$ such that:
%\begin{equation}
% \Gamma^{adv}_t= 100\frac {\bar{\tau}_{1974}.\exp(\gamma_t)} {\bar{\tau}_{19741}  \label{eq:tcadv_compoeffect}
%\end{equation}
%\noindent that is, Equation (\ref{eq:comp_effects_adv}) with $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean (ad-valorem) transport cost in 1974.

As for the additive component, given that the sector fixed effect and the country fixed effect are additive rather than multiplicative by construction, we estimate the following equation using non-linear least squares:\footnote{For sake of notational simplicity, we do not distinguish the coefficients associated with the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g. the series of $\gamma_t$ differs from one estimation to the other).
Note that we apply the same weighting scheme as for the OLS regression (based on the relative value of the flow).}

%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
\begin{eqnarray}
\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
\end{eqnarray}

As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)).
Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the goods dimension, while the additive cost is the sum of the two dimensions.
Both equations are estimated by transport mode.\smallskip

After estimating Equation (\ref{eq:compeffects_add}), we can rebuild the additive component according to:
\[\left\{
  \begin{array}{lcl}
\text{For the year 1974:}&&\widehat{t}_{is74}=  \delta + \alpha_i+ \beta_s, \\
\text{For any year}~t> 1974:&&\widehat{t}_{ist}=\left(\delta + \alpha_i+ \beta_s\right).\exp(\gamma_t)
\end{array}
\right.\]

From this, we deduce the recursive link: $\widehat{t}_{ist} = \widehat{t}_{is74} \times \exp(\gamma_t)$.
Given the constraint $t>0$, we then obtain the percentage change from 1974 from $\Gamma^{add}_{ist} = 100\frac{\widehat{t}_{ist}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)$. Again, it is independent of the product-origin country pair. We can thus rewrite the time-trend series for the additive transport cost component as:
\begin{equation}
\Gamma^{add}_t  = 100\exp(\gamma^{add}_t) \label{eq:tcadd_compoeffect}
\end{equation}

\noindent where the subscript $^{add}$ clarifies that this time fixed effect is specific to the additive component. As a result, the two series $\Gamma^{adv}_t$ and $\Gamma^{add}_t$ have a straightforward interpretation in percentage changes from the initial value of 100 for $t=1974$.

\paragraph{Obtaining the unfitted measures} - The objective here is to express the unfitted transport cost component (additive and multiplicative) built as an index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2019).
To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:

\textbf{ENLEVER LE TREND LIE A L'INFLATION. POURQUOI SEULEMENT ICI, ET PAS DANS LE FITTED? that exprime en USD aussi. REPRENDRE AVEC GUILLAUME}

@@

We also build two measures of the ``total'' transport costs, combining our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e. composition effects excluded). Specifically, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974. Even if obeying the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain.
\smallskip

\paragraph{Obtaining the unfitted total transport cost index} - For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}, \textbf{and $\widetilde{p}_t$ the corresponding average observed fas price}.
Recall that $\widehat{\tau}^{adv}_t$ measures the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed as percentage of the fas price.
We then transform this value in an index with base year 1974, applying a similar formula as above (by transport mode):
%$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$

$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t  }{\widehat{tc}^{raw}_{1974}}$$

\textbf{POURQUOI UN (-1) AVANT?}

@@

$$\widehat{\tau}^{pure, adv}_t = \bar{\tau}_{1974}\frac{\Gamma^{adv}_t }{100} $$

@@

with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost. \textbf{POURQUOI PAS tbar 1974? On reconstruit la mesure en USD? PAS CLAIR}

We then deduce the fitted value of the overall cost according to: \textbf{il manque le fas price, je l'ajoute pour la composante additive}
$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \frac{\widehat{t}^{pure}_t}{\widetilde{p}^{fas}_t}$$

@@

\paragraph{For the additive component} -
